---
title: "Seurat Analysis Results"
fontsize: 50pt
output: 
  html_document:
    theme: united
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = 'png')
```

```{r Initialization, message=FALSE, results='hide'}
# seurat_dir <- "."
source(sprintf("%s/visrutils.R", seurat_dir))

visr.app.start("Seurat", debugdata=mtcars, input.type = "none")
source(sprintf("%s/Utils.R",seurat_dir))
source(sprintf("%s/Seurat_IO.R",seurat_dir))
source(sprintf("%s/Seurat_filter.R",seurat_dir))
source(sprintf("%s/Seurat_steps.R",seurat_dir))
source(sprintf("%s/Seurat_find_var_genes.R",seurat_dir))
source(sprintf("%s/Seurat_dim_reduction.R",seurat_dir))
source(sprintf("%s/Seurat_cluster_cells.R",seurat_dir))
source(sprintf("%s/Seurat_DE.R",seurat_dir))
source(sprintf("%s/Seurat_additional.R",seurat_dir))
visr.app.end(printjson=F, writefile=F)
visr.applyParameters()

# I/O
visr.param.workflow <- 'single'
visr.param.Import_method <- 'load_raw'
visr.param.Path_to_outs <- input_dir # "C:/Users/Yiwei Zhao/Desktop/pbmc3k/"
visr.param.Output_directory <- output_dir # "C:/Users/Yiwei Zhao/Desktop/temp/"
visr.param.create_subdir <- F

print(visr.param.Path_to_outs)
print(visr.param.Output_directory)

#Filter/Create Seurat

#Seurat steps
visr.param.Find_Variable_Genes <- T
visr.param.Dim_Reduction <- T
visr.param.Cluster_Cells <- T
visr.param.Find_Marker_Genes <- T

#Find var genes

#Dim reduction
nPC_compute <- 20
visr.param.Run_PCA <- T
visr.param.nPC_compute <- nPC_compute
visr.param.jackstraw <- T
visr.param.nPC_jackstraw <- nPC_compute
visr.param.jackstrawPlot_ncol <- 5 # can be adjusted
visr.param.elbow <- T
visr.param.PC_heatmap <- T
visr.param.nPC_PCheatmap <- nPC_compute

visr.param.Run_tSNE <- T

#Cluster Cells

#DE
visr.param.Choose_Clusters <- "all"


# Check Input -------------------------------------------------------------
if (visr.param.workflow == "single"){
  if (visr.param.Import_method == "load_raw"){
    if (!dir.exists(visr.param.Path_to_outs)){
      visr.message(paste("'",visr.param.Path_to_outs,"'"," is not a valid path to Cell Ranger pipeline output directory. It should contain another directory named \"outs\""))
    }
  } else {
    if (!file.exists(visr.param.Path_to_seurat_object)){visr.message(sprintf("'%s' is not a valid path to seurat object", visr.param.Path_to_seurat_object))}
  }
}else{
  if (visr.param.Import_method2 == "load_one"){
    assert_that(file.exists(visr.param.seurat_integrated), msg = sprintf("'%s' is not a valid path to seurat object",visr.param.seurat_integrated))
  }else{
    assert_that(file.exists(visr.param.seurat_obj_1), msg = sprintf("'%s' is not a valid path to seurat object 1",visr.param.seurat_obj_1))
    assert_that(file.exists(visr.param.seurat_obj_2), msg = sprintf("'%s' is not a valid path to seurat object 2",visr.param.seurat_obj_2))
  }
}

if (!dir.exists(visr.param.Output_directory)){visr.message("Path to output directory not found")}

assert_that(visr.param.max_nGenes > visr.param.min_nGenes,msg = "The high cutoff of the number of genes must be greather than the low cutoff" )
assert_that(visr.param.Mean_exp_high > visr.param.Mean_exp_low, msg = "The high cutoff of the mean expression level must be greather than the low cutoff")
assert_that(visr.param.Dispersion_high > visr.param.Dispersion_low, msg =  "The high cutoff of dispersion must be greather than the low cutoff" )
# Load Packages -----------------------------------------------------------
# library("Rtsne",lib.loc = "/brcwork/bioinf/tools/R/R-3.4.3/lib64/R/library") #old version 0.13 didn't work
visr.library("Seurat")
visr.library("dplyr")
visr.library("reshape2")
visr.library("DT")

# Analysis result file names ----------------------------------------------

plot_output <- "plots.pdf"
cell_info_output <- "Cells.tsv"
DE_output <- "Differential_Expression_Analysis_result.tsv"
var_gene_output <- "Variable_Genes.tsv"
selected_gene_output <- "Selected_Marker_Genes.tsv"

# Create PDF --------------------------------------------------------------

# save plots to this location
if (visr.param.create_subdir){
  subfolder <- Sys.time()
  subfolder <- gsub(" ","_",subfolder)
  subfolder <- gsub(":","-",subfolder)
  subfolder <- gsub("-","",subfolder)
  output_folder <- paste(visr.param.Output_directory,subfolder,sep = "/")
  dir.create(output_folder)
} else {
  output_folder <- visr.param.Output_directory
}

#startReport(output_dir = output_folder)
gc()
```

## Introduction

This report shows the results of clustering and differential expression analysis using **Seurat**. The the script used to generate the report utilizes the script of the **Seurat** app in **VisR**.

The input data for this report can be found in ``r input_dir``.

For more details on **Seurat** R package, see <https://satijalab.org/seurat/>.

For more details on **VisR** platform and **Seurat** app, see <https://visrsoftware.github.io/>.

For addtional questions on the report, output tables or VisR, please contact Bernie Zhao (bernieyw@brc.ubc.ca).

## Filter Dataset
```{r Load Data, results='hide', out.width=400, eval = T}
# Load Data ---------------------------------------------------------------
min_fraction_of_cells <- visr.param.min_fraction_cells
min_number_of_genes <- visr.param.min_nGenes
gbmData <- load_data()

if (visr.param.workflow == "integrated"){
  groups <- levels(as.factor(gbmData@meta.data$group))
}

# Additional checkpoint ---------------------------------------------------

if (visr.param.workflow == "integrated" && visr.param.Import_method2 == "load_one"){
  assert_that(!is.null(gbmData@meta.data$group),msg = "Need to merge datasets first")
}

if (visr.param.workflow == "integrated" && visr.param.Import_method2 == "load_two"){
  assert_that(visr.param.dataset1_name != visr.param.dataset2_name, msg = "Label of the two datasets must be different.")
}

if (length(visr.param.export_gene_list) > 0){
  valid_gene_probes <- check_gene_list(gbmData,visr.param.export_gene_list)
}
gc()
```

The cellranger pipeline output contains `r ncol(gbmData@raw.data)` cells and `r nrow(gbmData@raw.data)` genes. After filtering, there are `r ncol(gbmData@data)` cells and `r nrow(gbmData@data)` genes.

## Find Variable Genes
```{r Find variable Genes, results='hide', eval = T}
gbmData <- find_variable_genes(gbmData)
gc()
```

The number of variable genes is `r length(gbmData@var.genes)`.

## PCA
`r nPC_compute` PCs are computed initially. Cells are plotted on PC1 vs PC2 in the figure below, colored by their library ids.
```{r PCA, results='hide', eval = T}
gbmData <- run_PCA(gbmData)
gc()
```


## Determine Significant PCs {.tabset}
The 3 methods below can be used to determine which PCs to use for downstream analysis.
```{r calculate number of PCs, results='hide',eval = T}
auto_num_PC <- calculate_nPC(gbmData)
gc()
```

### Elbow Method
```{r Elbow, results='hide',fig.width=6, eval = T}
gbmData <- draw_elbow(gbmData)
gc()
```

In this report, PC1 through PC`r auto_num_PC` are considered as significant.

### Jackstraw
```{r Jackstraw, results='hide', fig.height=10, fig.width=15, eval = T}
# gbmData <- run_jackstraw(gbmData)

print(paste("Running JackStraw"))
nPC <- min(visr.param.nPC_jackstraw,length(gbmData@dr$pca@sdev))
gbmData <- JackStraw(object = gbmData, num.replicate = visr.param.jackstrawRep, num.pc = nPC)
JackStrawPlot(object = gbmData, PCs = 1:nPC, nCol = visr.param.jackstrawPlot_ncol)
gc()

```

In this report, PC1 through PC`r auto_num_PC` are considered as significant.

### PC Heatmap
```{r PC Heatmap, results='hide', fig.width=15,fig.height=20, eval = T}
gbmData <- draw_PCHeatmap(gbmData)
gc()
```

In this report, PC1 through PC`r auto_num_PC` are considered as significant.

## t-SNE 
`r auto_num_PC` PCs are used for t-SNE. Cells are plotted on t-SNE projections in the figure below, colored by their library ids.

```{r t-SNE, results='hide', eval = T}
gbmData <- run_tSNE(gbmData)
gc()
```

## Clustering
`r auto_num_PC` PCs are used for clustering. Cells are plotted on t-SNE projections in the figure below, colored by their cluster ids.
```{r Clustering Cells, results='hide', eval = T}
gbmData <- cluster_cells(gbmData, "pca")
gc()
```

## Differential Expression Analysis {.tabset}

The heatmap below shows the expression of the top 2 DE genes (based on adjusted p-value) for each cluster. The table below shows the top 50 DE genes for each cluster. The complete list of DE genes can be found in the output file ``r DE_output``.

### Heatmap
```{r DE Analysis, results='hide', eval = T}
table <- diff_exp(gbmData)
gc()
```

### DE genes table
```{r table, echo=FALSE, eval = T}
visr.param.top_n <- 50
table <- table %>% group_by(cluster) %>% filter(row_number() <= visr.param.top_n)

table$cluster <- as.factor(table$cluster)
table[,c("pct_diff", "avg_logFC")] <- round(table[,c("pct_diff", "avg_logFC")],digits = 3)
table[,c("p_val", "p_val_adj")] <- round(table[,c("p_val", "p_val_adj")],digits = 5)
colnames(table)[1] <- "DE_gene_name"

#check if datatable is causing the high mem consumption in the rendering step
datatable(data = table, rownames = T, filter = list(position = 'top', clear = T))
```

## Output Files

```{r output, results='hide', eval = T}
export_var_genes(gbmData)
export_cells(gbmData)

# Reduce file size of saved object
if (!is.null(gbmData@calc.params$RunTSNE)){
  gbmData@calc.params$RunTSNE$... <- NULL
}

if (visr.param.save_obj){
  # save object
  print(paste("Saving object"))
  saveRDS(gbmData, file = paste(output_folder,"Seurat.rds",sep = "/"))
}

gc()
```

* `seurat_analysis_results.html`: A report containing data summary and output plots
* `Seurat.rds`: Seurat object that can be loaded to **VisR** or R scripts for customized analysis
* ``r var_gene_output``: Dispersion table of genes
* ``r cell_info_output``: Analysis results related to each cell
* ``r DE_output``: Differentially expressed genes for each cluster

```{r, results='hide'}
rm(gbmData)
rm(list = ls())
gc()
```


